---
title: "Cool"
author: "Robert Orttung"
date: "August 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages({
  library(leaflet)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(DT)
})
```

Load salmon data from the KNB

```{r}
data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/knb.92020.1"
# data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e"
esc <- tryCatch(
    read.csv("data/escapement.csv", stringsAsFactors = FALSE),
    error=function(cond) {
        message(paste("Escapement file does not seem to exist, so get it from the KNB."))
        esc <- read.csv(url(data_url, method = "libcurl"), stringsAsFactors = FALSE)
        return(esc)
    }
)

head(esc)

```

# prepare data set for calculating medians

```{r}
median_esc <- esc %>% 
  separate(sampleDate, c("Year", "Month", "Day"), sep = "-")
  
  head(median_esc)
```

```{r}
median_esc <- esc %>% 
  separate(sampleDate, c("Year", "Month", "Day"), sep = "-") %>% 
    group_by(Species, SASAP.Region, Year, Location) %>% 
  summarize(escapement = sum(DailyCount))
  
  head(median_esc)
```

```{r}
median_esc <- esc %>% 
  separate(sampleDate, c("Year", "Month", "Day"), sep = "-") %>% 
    group_by(Species, SASAP.Region, Year, Location) %>% 
  summarize(escapement = sum(DailyCount)) %>% 
    group_by(Species) %>% 
  summarize(median_escapement = median(escapement))

  head(median_esc)
```

Plot Escapement Values as a bar chart. geom col creates the columns and coord flip makes the bars horizontal. aes x and y


```{r}
ggplot(median_esc, aes(Species, median_escapement)) +
  geom_col() +
  coord_flip()

```

Make a map. We are starting from daily escapment data again and finding the unique locations at which escapment was measured. 

distinct finds all the distinct versions of a variable. drop all rows that don't have lats and longs. however, in practice we should fix up our dataset. 


```{r}
locations <- esc %>% 
  distinct(Location, Latitude, Longitude) %>% 
  drop_na()
datatable(locations)
```
          
          add tiles finds map on the internet

```{r}
leaflet(locations) %>% 
  addTiles() %>% 
  addMarkers(~ Longitude, ~ Latitude, popup = ~ Location)
```

Fix map so all locations are in Alaska as they should be

```{r}

locs <- locations %>% mutate(Longitude = abs(Longitude) * -1)

leaflet(locs) %>% 
  addTiles() %>% 
  addMarkers(~ Longitude, ~ Latitude, popup = ~ Location)
````



